---
title: "Preprocessing and Cleaning"
author: "Benzon Carlitos Salazar"
date: "October 22, 2022"
---

# Preprocessing and Cleaning

## Load and check the data

```{r setup}
library(magrittr)

train <- readr::read_tsv(here::here("data-raw", "drugsComTrain_raw.tsv"))
test <- readr::read_tsv(here::here("data-raw", "drugsComTest_raw.tsv"))

full <- 
  dplyr::bind_rows(train, test)

# Check our data
tibble::glimpse(full)

colnames(full)
```

Based on our output, we get `r as.integer(ncol(full))` variables (or columns) 
and `r as.integer(dplyr::count(full))` observations (or rows). That's a pretty 
lengthy dataset!

Our data dictionary is as follows:

| Variable Name | Description                             |
|:------------- |:--------------------------------------- |
| ...1          | Identifier                              |
| drugName      | Name of drug                            | 
| condition     | Name of condition                       |
| review        | Patient review                          |
| rating        | 10 star patient rating                  |
| date          | Date of review entry                    |
| usefulCount   | Number of users who found review useful |

## Tokenization and cleaning

We do some basic tokenization and cleaning.

The first column, `...1` looks to be an identifier, so we rename that column as 
and `id`, and then we do some minor tokenization to enforce consistency. 
Additionally, we will utilize the `{lubridate}` package to break down the date 
column to census month and census year, to see if it will give us some additional 
insights.

We will also remove any special characters in the `drugName` and `condition` 
columns, turn white spaces into underscores, while keeping the original text for 
the `review` for now.

```{r tokenization}
full <-
  full %>%
  dplyr::rename(id = "...1")

full$drugName <- stringr::str_replace_all(full$drugName, "[^[:alnum:]]", "_")
full$condition <- stringr::str_replace_all(full$condition, "[^[:alnum:]]", "_")
full$drugName <- tolower(full$drugName)
full$condition <- tolower(full$condition)
full$date <- lubridate::mdy(full$date)
full$rating <- as.numeric(full$rating)
full$usefulCount <- as.numeric(full$usefulCount)

full <-
  full %>%
  dplyr::mutate(census_month = lubridate::month(date)) %>%
  dplyr::mutate(census_year = as.character(lubridate::year(date))) %>%
  dplyr::rename(full_date = date)
```

We get a new looking dataset.

```{r new}
full %>%
dplyr::slice_head(n = 5) %>%
tibble::glimpse()
```

```{r save_and_clean, include=FALSE, echo=FALSE, warning=FALSE}
saveRDS(full, here::here("data", "full.rds"))

rm(list = ls())
gc()
```
