---
title: "Preprocessing and Cleaning"
author: "Benzon Carlitos Salazar"
date: "October 22, 2022"
---

# Preprocessing and Cleaning

## Load and check the data

```{r setup}
library(magrittr)

train <- readr::read_tsv(here::here("data-raw", "drugsComTrain_raw.tsv"))
test <- readr::read_tsv(here::here("data-raw", "drugsComTest_raw.tsv"))

full <- dplyr::bind_rows(train, test)

# Check our data
str(full)

colnames(full)
```

Based on our output, we get `r ncol(full)` variables (or columns) and 
`r dplyr::count(full)` observations (or rows). That's a pretty lengthy dataset!

Our data dictionary is as follows:

| Variable Name | Description                             |
|:------------- |:--------------------------------------- |
| ...1          | Identifier                              |
| drugName      | Name of drug                            | 
| condition     | Name of condition                       |
| review        | Patient review                          |
| rating        | 10 start patient rating                 |
| date          | Date of review entry                    |
| usefulCount   | Number of users who found review useful |

## Tokenization and cleaning

The first column, `...1` looks to be an identifier, so we rename that column as 
and `id`, and then we do some minor tokenization to enforce consistency. 
Additionally, we will utilize the `lubridate` package to break down the date 
column to census month and census year, to see if it will give us some additional 
insights.

We will also remove any special characters in the `drugName` and `condition` 
columns, turn white spaces into underscores, while keeping the original text for 
the `review` for now.

```{r tokenization}
full <-
  full %>%
  dplyr::rename(id = "...1")

full$drugName <- stringr::str_replace_all(full$drugName, "[^[:alnum:]]", "_")
full$condition <- stringr::str_replace_all(full$condition, "[^[:alnum:]]", "_")
full$drugName <- tolower(full$drugName)
full$condition <- tolower(full$condition)
full$date <- lubridate::mdy(full$date)
full$rating <- as.numeric(full$rating)
full$usefulCount <- as.numeric(full$usefulCount)

full <-
  full %>%
  dplyr::mutate(census_month = lubridate::month(date)) %>%
  dplyr::mutate(census_year = lubridate::year(date)) %>%
  dplyr::rename(full_date = date)
```

## Summary statistics

Doing some summarizations, we can see count distributions for each column. This 
will also help us determine if there exists any NA columns. For this analysis, 
we utilize the `gtsummary` package [(link)](https://www.danieldsjoberg.com/gtsummary/).

```{r summarization}
full %>%
gtsummary::tbl_summary()
```

## Patient review text cleaning

Let's take a look at 3 sample patients' review text to se what we are dealing 
with.

```{r pat_review_text}
sample_txt <-
  full %>%
  dplyr::select(id, review) %>%
  head(n = 5)
```

**Patient 1's review reads:** `r sample_txt[[2]][1]`

**Patient 2's review reads:** `r sample_txt[[2]][2]`

**Patient 3's review reads:** `r sample_txt[[2]][3]`

**Patient 4's review reads:** `r sample_txt[[2]][4]`

**Patient 5's review reads:** `r sample_txt[[2]][5]`

From this, we get the following initial observations:

* Capitalizations of names.
* We get a mix of numeric and character representation of numbers (i.e., 5 Mg vs 
two weeks).
* Word contractions -- it's, didn't, I'm, etc.
* Special characters -- dash (-), ellipsis (...), parentheses.



```{r save_and_clean, include=FALSE, echo=FALSE, warning=FALSE}
saveRDS(full, here::here("data", "full.rds"))

# rm(list = ls())
# gc()
```
