---
title: "Preprocessing and Cleaning"
author: "Benzon Carlitos Salazar"
date: "October 22, 2022"
---

# Preprocessing and Cleaning

## Load and check the data

```{r setup}
library(magrittr)

train <- readr::read_tsv(here::here("data-raw", "drugsComTrain_raw.tsv"))
test <- readr::read_tsv(here::here("data-raw", "drugsComTest_raw.tsv"))

full <- dplyr::bind_rows(train, test)

# Check our data
str(full)

colnames(full)
```

Based on our output, we get `r ncol(full)` variables (or columns) and 
`r dplyr::count(full)` observations (or rows). That's a pretty lengthy dataset!

Our data dictionary is as follows:

| Variable Name | Description                             |
|:------------- |:--------------------------------------- |
| ...1          | Identifier                              |
| drugName      | Name of drug                            | 
| condition     | Name of condition                       |
| review        | Patient review                          |
| rating        | 10 star patient rating                  |
| date          | Date of review entry                    |
| usefulCount   | Number of users who found review useful |

## Tokenization and cleaning

The first column, `...1` looks to be an identifier, so we rename that column as 
and `id`, and then we do some minor tokenization to enforce consistency. 
Additionally, we will utilize the `lubridate` package to break down the date 
column to census month and census year, to see if it will give us some additional 
insights.

We will also remove any special characters in the `drugName` and `condition` 
columns, turn white spaces into underscores, while keeping the original text for 
the `review` for now.

```{r tokenization}
full <-
  full %>%
  dplyr::rename(id = "...1")

full$drugName <- stringr::str_replace_all(full$drugName, "[^[:alnum:]]", "_")
full$condition <- stringr::str_replace_all(full$condition, "[^[:alnum:]]", "_")
full$drugName <- tolower(full$drugName)
full$condition <- tolower(full$condition)
full$date <- lubridate::mdy(full$date)
full$rating <- as.numeric(full$rating)
full$usefulCount <- as.numeric(full$usefulCount)

full <-
  full %>%
  dplyr::mutate(census_month = lubridate::month(date)) %>%
  dplyr::mutate(census_year = lubridate::year(date)) %>%
  dplyr::rename(full_date = date)
```

## Summary statistics

Doing some summarizations, we can see count distributions for each column. This 
will also help us determine if there exists any NA columns.

### Drugs

```{r drug_summarization}
drug_count <-
  full %>%
  dplyr::select(drugName) %>%
  dplyr::group_by(drugName) %>%
  dplyr::summarise(count = dplyr::n())
```

What is the top 10 most used drug?

```{r top_10_most_drug}
drug_count %>%
dplyr::arrange(desc(count)) %>%
dplyr::slice_head(n = 10) %>%
knitr::kable()
```

What is the top 10 least used drug?

```{r top_10_least_drug}
drug_count %>%
dplyr::arrange(count) %>%
dplyr::slice_head(n = 10) %>%
knitr::kable()
```

The top 10 least used drug is actually misleading, as there are 
`r drug_count %>% dplyr::filter(count == 1) %>% dplyr::count()` drugs that all 
have a count of 1.

Are there any NAs or NULLs for `drugName`?

```{r na_nulls}
drug_count %>%
dplyr::filter(is.na(drugName) | is.null(drugName)) %>%
dplyr::count()
```

### Condition

```{r condition_summarization}
condition_count <-
  full %>%
  dplyr::select(condition) %>%
  dplyr::group_by(condition) %>%
  dplyr::summarise(count = dplyr::n())
```

What are the top 10 patient conditions?

```{r pat_condition_top_10}
condition_count %>%
dplyr::arrange(desc(count)) %>%
dplyr::slice_head(n = 10) %>%
knitr::kable()
```

What are the patient conditions?

```{r pat_condition_least}
condition_count %>%
dplyr::arrange(count) %>%
dplyr::slice_head(n = 35) %>%
knitr::kable()
```

I am unsure what to make of the `*_span_users_found_this_comment_helpful_`, so 
we will be removing those later. But ignoring that weirdness, we can see that 
`amyotrophic_lateral_sclerosis` is the least mentioned condition, followed by 
`anti_nmda_receptor_encephalitis`.

### Ratings

Since the ratings are 1-10, we will categorize them into 3 levels:

* low_rating = 0-3
* mid_rating = 4-7
* high_rating = 8-10

First, we create our row levels, then we add a new `drug_rating` column.

```{r rating_row_lvls}
rating_row_lvl <- 
  c("Low [0-3]", 
    "Mid [4-7]", 
    "High [8-10]", 
    "Unknown")

full_proc_rating <-
  full %>%
  dplyr::mutate(
    drug_rating = factor(
      dplyr::case_when(
        rating > 0 & rating <= 3 ~ rating_row_lvl[1],
        rating >= 4 & rating <= 7 ~ rating_row_lvl[2],
        rating >= 8 & rating <= 10 ~ rating_row_lvl[3],
        TRUE ~ rating_row_lvl[4]
      ), levels = rating_row_lvl))
```

Then, we utilize `gtsummary` package to see count distribution and frequency 
distributions.

```{r rating_gt}
full_proc_rating %>%
dplyr::select(drug_rating) %>%
gtsummary::tbl_summary(label = list(drug_rating ~ "Drug Rating")) %>%
gtsummary::bold_labels()%>%
gtsummary::modify_footnote(update = dplyr::everything() ~ NA)
```

## Patient review text cleaning

Let's take a look at 3 sample patients' review text to se what we are dealing 
with.

```{r pat_review_text}
sample_txt <-
  full %>%
  dplyr::select(id, review) %>%
  head(n = 5)
```

**Patient 1's review reads:** `r sample_txt[[2]][1]`

**Patient 2's review reads:** `r sample_txt[[2]][2]`

**Patient 3's review reads:** `r sample_txt[[2]][3]`

**Patient 4's review reads:** `r sample_txt[[2]][4]`

**Patient 5's review reads:** `r sample_txt[[2]][5]`

From this, we get the following initial observations:

* Capitalizations of names.
* We get a mix of numeric and character representation of numbers (i.e., 5 Mg vs 
two weeks).
* Word contractions -- it's, didn't, I'm, etc.
* Special characters -- dash (-), ellipsis (...), parentheses.



```{r save_and_clean, include=FALSE, echo=FALSE, warning=FALSE}
saveRDS(full, here::here("data", "full.rds"))

# rm(list = ls())
# gc()
```
