---
title: "Summary statistics"
author: "Benzon Carlitos Salazar"
date: "October 22, 2022"
---

```{r setup, warning=FALSE, include=FALSE, echo=FALSE}
full <- readRDS(here::here("data", "full.rds"))
```

# Summary statistics

Doing some summarizations, we can see count distributions for each column. This 
will also help us determine if there exists any NA columns.

## Drugs

```{r drug_summarization}
drug_count <-
  full %>%
  dplyr::select(drugName) %>%
  dplyr::group_by(drugName) %>%
  dplyr::summarise(count = dplyr::n())
```

What is the top 10 most used drug?

```{r top_10_most_drug}
drug_count %>%
dplyr::arrange(desc(count)) %>%
dplyr::slice_head(n = 10) %>%
knitr::kable()
```

What is the top 10 least used drug?

```{r top_10_least_drug}
drug_count %>%
dplyr::arrange(count) %>%
dplyr::slice_head(n = 10) %>%
knitr::kable()
```

The top 10 least used drug is actually misleading, as there are 
`r drug_count %>% dplyr::filter(count == 1) %>% dplyr::count()` drugs that all 
have a count of 1.

Are there any NAs or NULLs for `drugName`?

```{r na_nulls}
drug_count %>%
dplyr::filter(is.na(drugName) | is.null(drugName)) %>%
dplyr::count()
```

## Patient Condition

```{r condition_summarization}
condition_count <-
  full %>%
  dplyr::select(condition) %>%
  dplyr::group_by(condition) %>%
  dplyr::summarise(count = dplyr::n())
```

What are the top 10 patient conditions?

```{r pat_condition_top_10}
condition_count %>%
dplyr::arrange(desc(count)) %>%
dplyr::slice_head(n = 10) %>%
knitr::kable()
```

What are the least patient conditions?

```{r pat_condition_least}
condition_count %>%
dplyr::arrange(count) %>%
dplyr::slice_head(n = 35) %>%
knitr::kable()
```

I am unsure what to make of the `*_span_users_found_this_comment_helpful_`, so 
we will be removing those later. But ignoring that weirdness, we can see that 
`amyotrophic_lateral_sclerosis` is the least mentioned condition, followed by 
`anti_nmda_receptor_encephalitis`.

## Ratings

Since the ratings are 1-10, we will categorize them into 3 levels:

* low_rating = 0-3
* mid_rating = 4-7
* high_rating = 8-10

First, we create our row levels, then we add a new `drug_rating` column.

```{r rating_row_lvls}
rating_row_lvl <- 
  c("Low [0-3]", 
    "Mid [4-7]", 
    "High [8-10]", 
    "Unknown")

full_proc_rating <-
  full %>%
  dplyr::mutate(
    drug_rating = factor(
      dplyr::case_when(
        rating > 0 & rating <= 3 ~ rating_row_lvl[1],
        rating >= 4 & rating <= 7 ~ rating_row_lvl[2],
        rating >= 8 & rating <= 10 ~ rating_row_lvl[3],
        TRUE ~ rating_row_lvl[4]
      ), levels = rating_row_lvl))
```

Then, we utilize `gtsummary` package to see count distribution and frequency 
distributions.

```{r rating_gt}
full_proc_rating %>%
dplyr::select(drug_rating) %>%
gtsummary::tbl_summary(label = list(drug_rating ~ "Drug Rating")) %>%
gtsummary::bold_labels() %>%
gtsummary::modify_footnote(update = dplyr::everything() ~ NA)
```

We see that the general consensus leans towards a high rating for each drugs. 
Let's drill down further to see what is the highest overall rating.

```{r rating_drill_down}
full %>%
dplyr::mutate(rating__char = as.character(rating)) %>%
dplyr::select(rating__char) %>%
gtsummary::tbl_summary(label = list(rating__char ~ "Drug Rating")) %>%
gtsummary::bold_labels() %>%
gtsummary::modify_footnote(update = dplyr::everything() ~ NA)
```

```{r tmp_rating, include=FALSE, echo=FALSE, warning=FALSE}
tmp_rating <-
  full %>%
  dplyr::mutate(rating__char = as.character(rating)) %>%
  dplyr::group_by(rating__char) %>%
  dplyr::summarise(count = dplyr::n())

max_tmp_cnt <- max(tmp_rating$count)
max_tmp_rating <-
  tmp_rating %>%
  dplyr::filter(count == max_tmp_cnt) %>%
  dplyr::select(rating__char)
```

We see that patients mostly rated their drugs `r max_tmp_rating`, with a maximum 
count of `r max_tmp_cnt`. This lets us know that the attitude towards each 
patient's drug is positive.

## Review Dates

Using the `census_month` and `census_year`, we should be able to see count 
distributions per month and per year. To make our analysis easier, we will 
summarize the census month and year separately.

First, we convert each month numerical representation to their correct string 
representation, and then summarize with `gtsummary`.

```{r review_months}
review_census_month <-
  full %>%
  dplyr::mutate(census_month__abbr = month.abb[census_month])

review_census_month %>%
dplyr::select(census_month__abbr) %>%
gtsummary::tbl_summary(
  label = list(
    census_month__abbr ~ "Census Month of Review"
  )
) %>%
gtsummary::bold_labels() %>%
gtsummary::modify_footnote(update = dplyr::everything() ~ NA)
```

```{r tmp_month, include=FALSE, echo=FALSE, warning=FALSE}
tmp__month <-
  review_census_month %>%
  dplyr::group_by(census_month__abbr) %>%
  dplyr::summarise(count = dplyr::n())

max_count <- max(tmp__month$count)

max_month <-
  tmp__month %>%
  dplyr::filter(count == max_count) %>%
  dplyr::select(census_month__abbr)
```

Based on the summary table, reviews look pretty evenly distributed throughout 
the months. Although, it looks like `r max_month` has the most reviews, with 
`r max_count` review counts.

Next, we take a look at the census year.

```{r review_years}
full %>%
dplyr::select(census_year) %>%
gtsummary::tbl_summary(
  label = list(census_year ~ "Census Year of Review")
) %>%
gtsummary::bold_labels() %>%
gtsummary::modify_footnote(update = dplyr::everything() ~ NA)
```

```{r tmp_year, include=FALSE, echo=FALSE, warning=FALSE}
tmp__yr <-
  full %>%
  dplyr::group_by(census_year) %>%
  dplyr::summarise(count = dplyr::n())

max_year_cnt <- max(tmp__yr$count)
max_year <- 
  tmp__yr %>%
  dplyr::filter(count == max_year_cnt) %>%
  dplyr::select(census_year)
```

At first glance, it looks like we see a linear increase in the number of patients 
review. We especially see a jump from 2014 to 2015, with year `r max_year` 
having the most reviews of `r max_year_cnt`.

```{r save_and_clean, include=FALSE, echo=FALSE, warning=FALSE}
rm(list = ls())
gc()
```
