---
title: "Patient Review"
author: "Benzon Carlitos Salazar"
date: "October 22, 2022"
---

```{r setup, warning=FALSE, include=FALSE, echo=FALSE}
library(magrittr)

full <- readRDS(here::here("data", "full.rds"))
```

## Patient review text cleaning

Let's take a look at 3 sample patients' review text to se what we are dealing 
with.

```{r pat_review_text}
sample_txt <-
  full %>%
  dplyr::select(id, review) %>%
  head(n = 5)
```

**Patient 1's review reads:** `r sample_txt[[2]][1]`

**Patient 2's review reads:** `r sample_txt[[2]][2]`

**Patient 3's review reads:** `r sample_txt[[2]][3]`

**Patient 4's review reads:** `r sample_txt[[2]][4]`

**Patient 5's review reads:** `r sample_txt[[2]][5]`

From this, we get the following initial observations:

* Capitalizations of names.
* We get a mix of numeric and character representation of numbers (i.e., 5 Mg vs 
two weeks).
* Word contractions -- it's, didn't, I'm, etc.
* Special characters -- dash (-), ellipsis (...), parentheses.

```{r save_and_clean, include=FALSE, echo=FALSE, warning=FALSE}
rm(list = ls())
gc()
```
