---
title: "Patient Review"
author: "Benzon Carlitos Salazar"
date: "October 22, 2022"
---

```{r setup, warning=FALSE, include=FALSE, echo=FALSE}
library(magrittr)

full <- readRDS(here::here("data", "full.rds"))
```

# Patient reviews

Let's take a look at 3 sample patients' review text to se what we are dealing 
with.

```{r pat_review_text}
sample_txt <-
  full %>%
  dplyr::select(id, review) %>%
  dplyr::slice_sample(n = 5)
```

**Patient 1's review reads:** `r sample_txt[[2]][1]`

**Patient 2's review reads:** `r sample_txt[[2]][2]`

**Patient 3's review reads:** `r sample_txt[[2]][3]`

**Patient 4's review reads:** `r sample_txt[[2]][4]`

**Patient 5's review reads:** `r sample_txt[[2]][5]`

From this, we get the following initial observations:

* Capitalization of names.
* We get a mix of numeric and character representation of numbers (i.e., 5 Mg vs 
two weeks).
* Word contractions -- it's, didn't, I'm, etc.
* Special characters -- dash (-), ellipsis (...), parentheses.
* Handle HTML character encodings.

## Basic Preprocessing

For some basic string cleaning, we can do the following:

* Remove trailing and leading whitespaces.
* Remove any newlines (`\n`) or tabs (`\t`).
* Normalize capitalizations.
* We convert HTML character encoding to their proper representations using 
`textutils` and enforce consistent character encoding -- `UTF-8`.
* Convert ellipsis

```{r sample_clean_up}
sample_txt_cleanup <- sample_txt[[2]][[3]]

cleaned_txt <-
  sample_txt_cleanup %>%
  textutils::HTMLdecode() %>%
  iconv("UTF-8", sub = "byte") %>%
  tolower() %>%
  stringr::str_replace_all("[^[:alnum:][.]['][!]]", " ") %>%
  stringr::str_trim() %>%
  stringr::str_squish() %>%
  stringr::str_replace_all("[.]{2,3}", ".") # convert ellipsis to a period
```

The following is an example of a cleaned text of patient 4's review:

**Original**: `r sample_txt_cleanup`

**Cleaned**: `r cleaned_txt`

```{r full_text_clean}
full$review <- 
  full$review %>%
  textutils::HTMLdecode() %>%
  iconv("UTF-8", sub = "byte") %>%
  tolower() %>%
  stringr::str_replace_all("[^[:alnum:][.]['][!]]", " ") %>%
  stringr::str_trim() %>%
  stringr::str_squish() %>%
  stringr::str_replace_all("[.]{2,3}", ".") # convert ellipsis to a period
```

```{r save_and_clean, include=FALSE, echo=FALSE, warning=FALSE}
saveRDS(full, here::here("data", "full_cleaned.rds"))

rm(list = ls())
gc()
```
