---
title: "Drug Reviews - Exploratory Data Analysis and Sentiment Analysis"
author: "Benzon Carlitos Salazar"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    code_folding: show
    df_print: paged
    toc: true
    toc_float: true
    theme: cosmo
  pdf_document:
    df_print: kable
bibliography: citations.bib
---

```{r setup, include=FALSE}
options(knitr.duplicate.label = "allow")
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
files <- list.files(pattern = "\\d+.*\\Rmd", recursive = TRUE)
chunks <- paste0('```{r child ="', files, '"}\n```\n')
cat(chunks, sep = "\n")
```

# Project Overview

Exploratory data analysis exercise of the drug review dataset that provides 
patient reviews on specific drugs along with related conditions and a 10 star 
patient rating reflecting overall patient satisfaction.

[Source Code](https://github.com/carrliitos/drug_reviews)

# Introduction

My goal for this project is to exercise my data analysis skills. I have chosen 
to work with the Drug Reviews dataset [@Gräßer2018] after spending some time in 
the UCI Machine Learning Repository [@Dua2019].

I will focus on some illustrative data visualizations, as well as, some word 
frequency analysis.

```{r child = "01_preprocessing.Rmd"}
```

```{r child = "02_summary_statistics.Rmd"}
```

```{r child = "03_patient_review.Rmd"}
```

```{r child = "04_corpus_statistics.Rmd"}
```

# Conclusion

I think this is where we will finish this project. Based on our corpus statistic, 
out of the 1,000 reviews we selected per year, we can see that `take` is the most 
frequent word used. And this makes sense because our corpus is on drug reviews, 
and the act of ***taking*** a drug is probably the only thing you do with 
medicine. Unless, of course the medicine was taken as a form of injection like 
a vaccination. However, it still makes sense because these are reviews of 
patients of drugs that they administered themselves.

All in all, if I had more computing power, perhaps I can do this on the full 
data set. Perhaps in the future, the computation itself can be done in the cloud, 
as well as any storage I may need.

# Packages Used

```{r packages_used, echo=FALSE, include=FALSE}
pckgs <-
  renv::dependencies() %>%
  tibble::as_tibble()

installed <-
  installed.packages() %>%
  tibble::as_tibble()

used <-
  pckgs %>%
  dplyr::inner_join(installed, by = "Package", suffix = c(".renv", ".base")) %>%
  dplyr::distinct(Package, Version.base) %>%
  dplyr::arrange(Package) %>%
  dplyr::rename(Version = "Version.base") %>%
  knitr::kable()
```

`r used`

# References
