---
title: "Drug Reviews - EDA"
author: "Benzon Carlitos Salazar"
date: "October 22, 2022"
output: 
  html_notebook:
    code_folding: show
    df_print: paged
    toc: TRUE
    toc_float: TRUE
    theme: cosmo
  pdf_document:
    df_print: kable
bibliography: citations.bib
---

```{r setup, include=FALSE}
options(knitr.duplicate.label = "allow")
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
files <- list.files(pattern = "\\d+.*\\Rmd", recursive = TRUE)
chunks <- paste0('```{r child ="', files, '"}\n```\n')
cat(chunks, sep = "\n")
```

# Project Overview

Exploratory data analysis exercise of the drug review dataset that provides 
patient reviews on specific drugs along with related conditions and a 10 star 
patient rating reflecting overall patient satisfaction.

# Introduction

My goal for this project is to exercise my data analysis skills. I have chosen 
to work with the Drug Reviews dataset [@Gräßer2018] after spending some time in 
the UCI Machine Learning Repository [@Dua2019].

I will focus on some illustrative data visualizations, and I will also try to 
stay true to the original paper and do a very basic, beginner-friendly sentiment 
analysis of drug satisfaction.

```{r child = "01_preprocessing.Rmd"}
```

```{r child = "02_summary_statistics.Rmd"}
```

```{r child = "03_patient_review.Rmd"}
```

# References
